\titlebox[english]


\Problem

Consider the function
%
\begin{equation}
    f(u, v) = u^2 + uv + 3v^2 + u^4
\end{equation}
%
on $\R^2$. 

\begin{subproblem}
    The function has a single stationary point $\vx_{*}$. Calculate it, and establish that it is indeed a local minimum point.
\end{subproblem}

\begin{subproblem}
    \begin{subsubproblem}
        Use $\vx_0 = (1, 1)$ as the initial point for a gradient search with constant step size $\alpha = \num{0.1}$ to find $x_{*}$, and calculate the iterate $\vx_1$.
    \end{subsubproblem}
    %
\begin{subsubproblem}
    Use $\widetilde{\vx}_0 = (1, 1)$ as the initial point for a Newton search to find $x_{*}$, and calculate the iterate $\widetilde{\vx}_1$.
\end{subsubproblem}
\end{subproblem}

\begin{subproblem}
    Consider any $C^2$ function $\fun[h]{\R^n}{\R}$. 
    It is well know that the Newton step is not always descent direction.
    Assuming that $\vx_k$ is not a stationary point of $h$, show that 
    it is possible to choose $\alpha_k \geq 0$ such that
    %
    \begin{equation}
        \vq_k = \vp_k - \alpha_k (\nabla h) (x_k),
    \end{equation}
    %
    is a descent direction for $h$ at $\vx_k$, where $\vp_k$ is the Newton step for $h$ at $\vx_k$.
\end{subproblem}


\newpageNotLF

\Problem 

(Counts as two questions) Each day a hospital requires the following number of nurses' aides (hjelpepleiere) on duty:

\begin{table}[H]
    \centering
    \caption{}
    \label{tab:my_label}
    \begin{tabular}{r@{}c@{}r|S[table-format=2.0]}
    \toprule 
    \multicolumn{3}{c}{Clock period} & {Minimum number on duty} \\
    \midrule
        7\AM &\ - \ & 11\AM & 60 \\
       11\AM &  -   &  3\PM & 70 \\
        3\PM &  -   &  7\PM & 60 \\
        7\PM &  -   & 11\PM & 50 \\
       11\PM &  -   &  3\AM & 20 \\
        3\AM &  -   &  7\AM & 30 \\
    \bottomrule
    \end{tabular}
\end{table}

Nurses' aides always arrive at the beginning of a clock period and work for
$8$~hours. Except that no nurses' aides arrive at $3\AM$ The salary per hour is $\SI{50}{\percent}$ higher from $7\PM$ to $7\AM$. Set up a linear programming model for a roster (bemanningsplan) of nurses' aides that ensures that the minium number required s always on duty, and that minimises the salary cost.

Do not solve the model, and ignore that in principle the model should be set up as an ILP problem.

\Problem

Consider the variational problem
%
\begin{equation}
    \int_0^1 \e^x \bigl[ 2 y(x)^2 + y'(x)^2 \bigr] \dx \to \min
\end{equation}
%
with $y(0)=0$, $y(1) = \e + \e^{-2}$ for $y \in C^2[0, 1]$.

\begin{subproblem}
    Find the Euler-Lagrange equation for the problem.
\end{subproblem}

\begin{subproblem}
    Establish a minimizer for the problem.
\end{subproblem}

\begin{subproblem}
    Find a valid candidate for an optimal solution of the extremal problem
    %
    \begin{align}
        \int_0^1 \e^x \bigl[ 2 y(x)^2 + y'(x)^2 \bigr] \dx 
        & \to \min, \\
        \int_0^1 e^x y(x) \dx & = 0,
    \end{align}
    %
    with $y(0) = 0$, $y(1) = 1$ for $y \in C^2[0,1]$. You are not asked to eastablish that it is a minimzer.
\end{subproblem}


\Problem

\begin{subproblem}
    \label{subproblem:MAT-2202-03-05-2015-4a}
    The linear program
    %
    \begin{align*}
        \phantom{2}x_1 + 2 x_2 + 4x_3 \to &\min \\
        4x_1 + \phantom{2}x_2 + 2x_3 \geq{}&{}3 \\
        2x_1 + 4x_2 + \phantom{4}x_3 \geq{}&{}5 \\
        x_1, x_2,x_3 \geq & \ 0 
    \end{align*}
    %
    is in canonical form. Write down its dual linear program, and solve the dual.
    (Find an optimal solution and the corresponding optimal value).
\end{subproblem}

\begin{subproblem}
    Writing down the linear program in \cref{subproblem:MAT-2202-03-05-2015-4a} in standard form
    %
    \begin{alignat*}{2}
        \phantom{2}x_1 + 2 x_2 + 4x_3 && \to &\min \\
        4x_1 + \phantom{2}x_2 + 2x_3 &&{}- e_1 \geq & \ 3 \\
        2x_1 + 4x_2 + \phantom{4}x_3 &&{}- e_2 \geq & \ 5 \\
        x_1, x_2,x_3,e_1,e_2 &&  \geq & \ 0 
    \end{alignat*}
    %
    Check that $x = \mat{0&1&1&0&0}\tran$ is a basic feasible solution. Establish that it is not an optimal solution.
\end{subproblem}